---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# json2tools

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/jsontools)](https://CRAN.R-project.org/package=jsontools)
<!-- badges: end -->

## Overview
With the increasing support of JSON in databases and since 2016 even support 
in the SQL standard, JSON data in vectors become more common. With jsontools
one can easily work with JSON vectors in R. 
The three main parts are:
1. parsing from JSON and converting to JSON
2. direct manipulation of JSON without parsing and conversion over R
3. extract data from JSON without parsing everything

## Installation

jsontools is not yet release on CRAN. You can install it from GitHub with

``` r
# install.packages("devtools")
devtools::install_github("tidyverse/dplyr")
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(jsontools)

gh_orgs[1]

gh_orgs <- prettify(gh_orgs)
gh_orgs[1]

json_glimpse(gh_orgs[1])

gh_orgs_small <- gh_orgs %>% 
  json_build_object(login, id, description)

parse_json_vector(gh_orgs_small, simplifyDataFrame = TRUE)

(logins <- jsonr_extract_chr(gh_orgs_small, ".login"))

json_mutate(gh_orgs, .login = paste0("new-", logins))[1:2] %>% 
  prettify()
```
