---
title: "Reading and Writing JSON"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Reading and Writing JSON}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(jsontools)
```

## Parsing from JSON

There are already a couple of packages to parse JSON (e.g. `rjson` or `jsonlite`). But for me they all had some drawbacks when working with JSON columns from a database:

* **bad/no vectorisation**: `RJSONIO` and `rjson` silently use only the first element of the input; `jsonlite` collapses the input, producing invalid JSON and then complains about trailing garbage.
* **handling of missing values**: `NA` and `NULL` input produce unclear errors and one cannot control how to handle them.
* **syntax**: the camel case syntax doesn't fit to the rest of the tidyverse.

`jsontools` relies on [`jsonlite`](https://cran.r-project.org/package=jsonlite) for parsing from JSON and conversion to JSON. For parsing it has the following functions:

* `parse_json()` to parse a character vector of length 1 (i.e. a scalar).
* `parse_json_vector()` to parse a character vector of any length.

As an example assume we got the following JSON vector

```{r}
customer_infos <- json2(c(
  '{"name": "Peter", "age": 19, "premium": true}',
  '{"name": "Daniel", "age": 41}',
  NA,
  '{"name": "Pablo", "age": 27, "premium": false}'
))

customer_infos
```

By default the `NA` produces an error

```{r error=TRUE}
parse_json_vector(customer_infos)
```

but with the `.na` argument we can specify which value should be used instead:

```{r}
parse_json_vector(customer_infos, .na = NULL)
```

Note that the default for `simplifyDataFrame` and `simplifyMatrix` has been changed to `FALSE`.
While they can be quite nice for interactive parsing of JSON they do not allow you to control how the result should look like.
For a more controlled conversion of the resulting list to a data frame have a look at the [`tibblify package`](https://cran.r-project.org/package=tibblify).


## Conversion to JSON

* **no vectorised version**: One cannot easily convert every element of a vector/list to JSON and get back a JSON vector.
* **convert every row of a data frame**: it is not so easy and/or slow to convert every row of a data frame into JSON.

For conversion there are the following functions:

* `format_json()`: convert an R object to JSON.
* `format_json_vector()`: convert each element of an R object to JSON and return them as a JSON vector.
* `format_json_rowwise()`: convert each row of a data frame to JSON.
* `write_json()`: convert an R object to JSON and write it to disk.

Let's look at our customer infos from above. Assume we have converted it into a tibble:

```{r}
customer_infos_df <- tibble::tibble(
  name = c("Peter", "Daniel", NA, "Pablo"),
  age = c(19L, 41L, NA, 27L),
  premium = c(TRUE, NA, NA, FALSE)
)
```

and filled the missing fields:

```{r}
customer_infos_df[3, ] <- tibble::tibble(
  name = "Michael",
  age = 38,
  premium = FALSE
)
customer_infos_df$premium[2] <- TRUE

customer_infos_df
```

We can now easily convert it back into a JSON vector

```{r}
format_json_rowwise(customer_infos_df)
```
